### 2015-04-23 manual/tex/advanced.tex
* started with logger/tracer manual

\section{Time}
\paragraph{Explain default time implementation}
\paragraph{Point to where the typedef can be changed}
\paragraph{Explain how users can create their own whacky implementation of the Time class for if the current version does not suit their needs}

\section{Tracing}
\paragraph{Briefly explain where the user can change the tracerset typedef}
\paragraph{Explain individual tracers}
\paragraph{Explain how the user can implement their own Tracer or OutputPolicy}