### 2015-11-10 paper/abstract.tex

With the ever increasing complexity of simulation models, parallel simulation
becomes necessary to perform the simulation within reasonable time bounds.
The built-in parallelism of Parallel DEVS is often insufficient to tackle this
problem on its own. Several synchronization algorithms have been proposed,
each with a specific kind of simulation model in mind. Due to the significant
differences between these algorithms, current Parallel DEVS simulation tools
restrict themself to only one such algorithm. In this paper, we present a
Parallel DEVS simulator, grafted on C++11, which offers both conservative and
optimistic simulation. We evaluate the performance gain that can be obtained
by choosing the most appropriate synchronization protocol. Our implementation
is compared to ADEVS using hardware-level profiling on a spectrum of benchmarks.