### 2015-11-29 paper/5-related-work.tex
Did some reading. Tried to fix some stuff we discussed in issue #28. Also changed the sentence structure in some places.

%TODO explain about other DEVS simulation kernels, and how they differ
%TODO explain about PythonPDEVS, the most related in terms of feature and "legacy" --> now in C++11 and way better performance
\subsection{PythonPDEVS}
Dxex is closely related to PyPDEVS in design and philosophy. PyPDEVS allows anyone who grasps the DEVS formalisms to immediately simulate his/her model without having to consider the kernel implementation. C++ implementations cannot hope to match the fast prototype/edit/run cycles provided by PyPDEVS, although this can be minimized by building the kernels as libraries. %maybe warn that if a header is touched, you still have to compile from scratch ? (eg. TIME_FP)
Advanced features such as activity based relocation and the performance gains this results in, are still unique to PyPDEVS. \\
\subsection{Adevs}
%TODO explain about adevs, the most related in terms of performance --> we implement more synchronization options and offer better performance because ...
Adevs's source code is still under active development, allowing for an exact comparison in performance and features. It remains in most aspects the fastest simulation engine for the DEVS formalism, but it lacks an optimistic synchronization implementation. %Tracing ?
By virtue of not flattening Coupled Models, performance suffers in increasingly hierarchical models.
%checkme
\subsection{CD++}
%TODO explain about CD++, the most related in terms of synchronization --> we implement everything in a single program, with a non-fragmented code-base
%TODO make mention of Warped, which is the kernel used in CD++, and why we don't use it
Different projects on CD++ offer conservative (CCD++) as well as optimistic (PCD++) parallel simulation. In contrast to our single program, with a non-fragmented code-base, neither projects offer both synchronization protocols. CD++ relies on the WARPED kernel. It is a middleware that provides memory, event, file, time and communication scheduling. WARPED is not used here since we operate explicitly on a shared memory system and since we wanted to design our kernels using the least amount of overhead possible.