### 2016-06-24 publications/Journal/conclusion.tex

\subsection{Conclusion}
We have shown that our contribution is invaluable for high performance simulation: depending on the expected behaviour, modellers can choose the most appropriate synchronization protocol.
We will summarize the key aspects in achieving good parallel performance here. % reword aspects syn
\subsubsection{Allocation}
Allocation of atomic models over kernels is critical to obtain a speedup. In the case of the Interconnect model it is not possible to allocate models without introducing a cyclic dependency between kernels with severe performance degradation as a result. 
Even when no cycles exist in the topology between kernels a good allocation scheme will minimize the number of connections between kernels and aim for a star topology as in PHoldTree with depth first allocation or a chain topology as in the Queue model. 
Dxex can optionally offer the user a visualization of the behavior of the kernels under a user supplied allocation scheme to allow for more insight. 
The same instrumentation could in future be used to optimize an existing allocation scheme.
\subsubsection{Uncertainty}
A simulation where future behaviour cannot be predicted is typically not suited for conservative simulation, but we have demonstrated that dxex's conservative kernel can still offer a non trivial speedup in this context. 
Optimistic is not constrained by this uncertainty and can offer a good speedup regardless of lookahead. 
PHOLD and PholdTree demonstrate that a single model can benefit from different synchronization protocols depending on a single parameter.
\subsubsection{Limits}
Dxex's conservative kernel is constrained by the amount of models it controls, as shown in the PHoldTree benchmark. This is especially the case when the minimal lookahead is $\epsilon$ requiring near constant polling of models for lookahead values. 
Optimistic is not constrained by the number of models it manages, but can quickly consume all available memory in a simulation. While this effect can be reduced with a thread aware allocation scheme, the underlying problem remains.