### 2016-06-28 publications/Journal/various.tex

As we will demonstrate in ~\nameref{sec:4-subsec:overhead-pgraph:memory} memory allocation is a performance bottleneck in simulations with high event frequency, especially in parallel with optimistic synchronization. Dxex has a modular allocation interface that separates the object creation in kernels and models from the actual implementation. This allows us by using a single compilation parameter to change the allocation strategy. This interface allows instrumentation of memory usage. Not only useful to detect memory leaks should this be a concern, it can also be of use to a model author to trace event creation in large models. The default implementation is the fastest common configuration verified by the benchmarks in this publication.