### 2016-09-30 publications/Journal/4-performance.tex
Initial section of pdevs.
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

% DESIGN OF MODELS
\input{models.tex}

% SEQUENTIAL
\input{sequential-performance.tex}

% PARALLEL
\input{parallel-performance.tex}

\input{pdevs.tex}

\input{memory.tex}

% CONCLUSION
\input{conclusion.tex}