### 2016-12-08 publications/Journal/8-conclusions.tex

In this paper, we introduced \textit{DEVS-Ex-Machina} (``\textit{dxex}''), a new C++11-based \textsf{Parallel DEVS} simulation tool.
Our main contribution is the implementation of multiple synchronization protocols for parallel simulation.
We have shown that there are indeed models which can be simulated significantly faster using either synchronization protocol.
\textit{Dxex} allows the user to choose between any of the six supported synchronization protocols.
We distinguish between inter-kernel (none, conservative, and optimistic synchronization) and intra-kernel (sequential or parallel concurrent transitions) synchronization protocols, which are orthogonal to one another.
Depending on observed model behaviour and simulation performance, runtime switching between synchronization protocols can be used.

Notwithstanding our modularity, \textit{dxex} achieves performance competitive to \textit{adevs}, another very efficient \textsf{DEVS} simulation tool.
Performance is measured both in elapsed time, and memory usage.
Our empirical analysis shows that allocation of models over kernels is critical to enable a parallel speedup.
Furthermore we have shown when and why optimistic synchronization can outperform conservative and vice versa.
We have also shown the influence of using the parallelism inherent in the Parallel DEVS formalism, and its interaction with inter-kernel synchronization algorithms.
Finally we investigated the effect of memory (de)allocation on parallel simulation.

Future work is possible in several directions.
First, our implementation of optimistic synchronization should be more tolerant to low-memory situations.
In its current state, simulation will simply halt with an out-of-memory error.
Having simulation control, which can throttle the speed of nodes that use up too much memory, has been shown to work in these situations~\cite{FujimotoBook}.
Faster GVT implementations~\cite{Fujimoto:1997:CGV:268403.268404,Bauer:2005:SND:1069810.1070159} might further help to minimize this problem.
Second, the runtime switching between synchronization protocols can be driven using machine learning techniques.
The simulation engine is already capable of collecting data to inform such a process, and is designed to listen for commands from an external component.
Third, automatic allocation might be possible by analysis of the connections between models.
This information is already used in \textit{dxex} to construct the dependency graph in conservative synchronization.
A graph algorithm that distributes models, while avoiding cycles, could be used to offer a parallel speedup in either optimistic or conservative synchronization.
Similarly, it could serve as a default parallel allocation scheme that can be improved by the user.
Finally, the use of transactional memory can offer several advantages in this project.
If it becomes part of the new C++17 standard it would be of great interest to see if it can help reduce the performance effects of memory allocation and synchronization.